\prefacesection{Acknowledgements}

Most importantly, I want to thank my advisor, Susan Holmes. She offered
motivation and valuable perspective for all the projects discussed in this
thesis, and I look to her for examples of how a true scholar works and thinks.
Thank you Professors Paul Switzer, Mike Baiocchi, Brad Efron, and David Relman
for serving on the thesis committee and offering useful feedback on my research.
Thank you to Lester Mackey and Persi Diaconis for shaping some of my basic views
of statistics and its role in the world.

Thank you to the Holmes lab -- Claire Donnat, Lan Huong Nguyen, Julia Fukuyama,
Christof Seiler, Pratheepa Jeganathan, Nikos Ignatiades, Ben Callahan, Sergio
Bacallado, Joey McMurdie -- for creating a welcoming and intellectually
stimulating environment. Thank you project collaborators -- Les Dethlefsen, Lina
Qiu, Suzanne Tamang, Shuo Xie, Yan Min -- for the creative ideas and fun
<<<<<<< HEAD
discussions. Thank you friends -- Joey Arthur, Jessy Hwang, Subhabrata Sen, Zhou
=======
discussions. Thank you friends -- Joey Arthur, Jessy Huang, Subhabrata Sen, Zhou
>>>>>>> fd4ab86d528aca46fe3ba122f48ebaad65253166
Fan, Mona Azadkia, Stephen Bates, Keli Liu, Pragya Sur, Gene Katsevich, Robin
Jia, Wanning Chen, Chris Rogers, Ling Li, Prof. Timothy Zerlang, ... -- for
making life here so interesting.

And finally thank you to my family for teaching me everything that was
important, besides just statistics.

I appreciate the funding my work has received through an NIH training grant (T32
GM 96982-5) and a Stanford University Weiland fellowship.
